---
title: "Plotly"
author: "Hannah Bougdah & Alex Ma"
date: "29 July 2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## coffee and coding: Plotly

Today's presentation

- What is Plotly?

- Online resources

- example code using ggplot as a basis

- example code using plotly



## Plotly

- Plotly is a graphing library which makes interactive html graphs.

- It uses the open source JavaScript graphing library **plotly.js**.

- Great for building dashboards or allowing the user to interact with the data themselves.



## Getting started

- Graphing library with example code: https://plot.ly/r/

- Cheat sheet (paste into web browser): https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf

- Book is available too: https://plotly-r.com/index.html

- most data used in this presentation is available here:
https://data.gov.uk/dataset/cb7ae6f0-4be6-4935-9277-47e5ce24a11f/road-safety-data


## Functionality

```{r, include = FALSE}
# create dft colour palette

library(fBasics)

r <- c(0, 102, 210, 228, 201, 233, 0, 153)
g <- c(104, 164, 95, 159, 146, 211, 153, 214)
b <- c(83, 152, 21, 115, 18, 160, 169, 221)

dft_col <- function (n, name = c("dft")) 
{
  dft = rgb(r, g, b, maxColorValue = 255)
  name = match.arg(name)
  orig = eval(parse(text = name))
  rgb = t(col2rgb(orig))
  temp = matrix(NA, ncol = 3, nrow = n)
  x = seq(0, 1, , length(orig))
  xg = seq(0, 1, , n)
  for (k in 1:3) {
    hold = spline(x, rgb[, k], n = n)$y
    hold[hold < 0] = 0
    hold[hold > 255] = 255
    temp[, k] = round(hold)
  }
  palette = rgb(temp[, 1], temp[, 2], temp[, 3], maxColorValue = 255)
  palette
}

```

```{r libraries, include=FALSE}
library(dplyr)
library(lubridate)
library(readxl)
library(plotly)


# read in data
accidents <- read.csv("Acc.csv")


# need to import casualties as excel due to weird scientific notation
casualties <- read_excel("cas.xlsx", col_types = c("text", 
    "numeric", "numeric", "numeric", "numeric", 
    "numeric", "numeric", "numeric", "numeric", 
    "numeric", "numeric", "numeric", "numeric", 
    "numeric", "numeric", "numeric"))


# rename accident index
accidents <- dplyr::rename(accidents, Accident_Index = "ï..Accident_Index")
#accidents <- dplyr::rename(accidents, Accident_Index = "?..Accident_Index")

# make accident index factor not character
accidents$Accident_Index <- as.character(accidents$Accident_Index)



# link two date from accident onto casualties data set
casualties <- casualties %>%
  dplyr::left_join(select(accidents, Date, Accident_Index), by = "Accident_Index")

############# create time-series by month

# prepare month variable
crashes_dates <- casualties %>%
  group_by(month = lubridate::month(Date))

# group data, count how many casualties in each casualty and remove missing
crashes_dates <- crashes_dates %>%
  dplyr::group_by(month, Casualty_Type) %>%
  dplyr::tally() %>%
  dplyr::filter(Casualty_Type != -1)


crashes_dates <- crashes_dates %>%
  dplyr::filter(!is.na(month))

# make character otherwise look up won't recognise 0 in casualty type
crashes_dates$Casualty_Type <- as.character(crashes_dates$Casualty_Type)
                
                
# create casualty type look up using look up from data.gov
lookup_cas <- c("0" = "Pedestrian", "1" = "Pedal Cyclist", "2" = "Motorcycle", "3" = "Motorcycle", "4" = "Motorcycle", 
"5" = "Motorcycle", "23" = "Motorcycle", "97" = "Motorcycle", "8" = "Car", "9" = "Car", "10" = "Car", "11" = "Bus or Coach", 
"19" = "Vans or Goods", "20" = "HGV", "21" = "HGV", "16" = "Other", "17" = "Other", "18" = "Other", "22" = "Other",
"90" = "Other", "98" = "Other")

# assign lookup to casualty type variable
crashes_dates$Casualty_Type <- lookup_cas[crashes_dates$Casualty_Type]

# group by casualty type
crashes_dates <- crashes_dates %>%
  dplyr::group_by(month, Casualty_Type) %>%
  dplyr::summarise(n = sum(n))


# ggplot + plotly chart - month
p_date <- ggplot(crashes_dates, aes(month, n))+
  geom_line(aes(colour = Casualty_Type), size = 1) +
  ylab("Casualties per month") +
  theme(axis.title = element_text(family = "Arial", size = 7)) +
  scale_x_continuous(breaks = round(seq(min(crashes_dates$month), max(crashes_dates$month), by = 1.0),1)) +
  scale_y_continuous(breaks = seq(0, 9000, by = 1000)) +
  scale_color_manual(values=dft_col(n=8, name="dft")) +
  theme_classic()



```

```{r plot, echo = FALSE, message = FALSE}

ggplotly(p_date)

```

## Why use Plotly?

- Good for exploring the data.
- Allow other users to explore the data.
- Easily save and download the charts you create in plotly.
- Not neccessarily good for publication - the charts are html.
- Useful in exploratory analysis or QA notes.


## Using plotly

There are two ways to use plotly:

- the easier way (less options)
- the harder way (more options)


## The easier way

The easier way involves creating charts using **ggplot2** and then using the **ggplotly** function from plotly to make the ggplot chart interactive. 

This is a quick, easy way to create interactive charts, however loses some of the functionality of the plotly package.


## The easier way - example 1




```{r code example 1, echo = TRUE, message = FALSE}

# ggplot + plotly chart - month
p_date <- ggplot(crashes_dates, aes(month, n))+
  geom_line(aes(colour = Casualty_Type), size = 1) +
  ylab("Casualties per month") +
  theme(axis.title = element_text(family = "Arial", size = 7)) +
  scale_x_continuous(breaks = round(seq(min(crashes_dates$month), max(crashes_dates$month), by = 1.0),1)) +
  scale_y_continuous(breaks = seq(0, 9000, by = 1000)) +
  scale_color_manual(values=dft_col(n=8, name="dft")) +
theme_classic()

ggplotly(p_date)

```


## The easier way - example 2

```{r code example 2, include = FALSE, message = FALSE}


#corresponding values are found in the data.gov. link in the "variable lookup" under supporting documents
accidents$Police_Force[accidents$Police_Force ==1] <- "Metropolitan Police"
accidents$Police_Force[accidents$Police_Force ==3] <- "Cumbria"
accidents$Police_Force[accidents$Police_Force ==4] <- "Lancashire"
accidents$Police_Force[accidents$Police_Force ==5] <- "Merseyside"
accidents$Police_Force[accidents$Police_Force ==6] <- "Greater Manchester"
accidents$Police_Force[accidents$Police_Force ==7] <- "Cheshire"
accidents$Police_Force[accidents$Police_Force ==10] <- "Northumbria"
accidents$Police_Force[accidents$Police_Force ==11] <- "Durham"
accidents$Police_Force[accidents$Police_Force ==12] <- "North Yorkshire"
accidents$Police_Force[accidents$Police_Force ==13] <- "West Yorkshire"
accidents$Police_Force[accidents$Police_Force ==14] <- "South Yorkshire"
accidents$Police_Force[accidents$Police_Force ==16] <- "Humberside"
accidents$Police_Force[accidents$Police_Force ==17] <- "Cleveland"
accidents$Police_Force[accidents$Police_Force ==20] <- "West Midlands"
accidents$Police_Force[accidents$Police_Force ==21] <- "Staffordshire"
accidents$Police_Force[accidents$Police_Force ==22] <- "West Mercia"
accidents$Police_Force[accidents$Police_Force ==23] <- "Warwickshire"
accidents$Police_Force[accidents$Police_Force ==30] <- "Derbyshire"
accidents$Police_Force[accidents$Police_Force ==31] <- "Nottinghamshire"
accidents$Police_Force[accidents$Police_Force ==32] <- "Lincolnshire"
accidents$Police_Force[accidents$Police_Force ==33] <- "Leicestershire"
accidents$Police_Force[accidents$Police_Force ==34] <- "Northamptonshire"
accidents$Police_Force[accidents$Police_Force ==35] <- "Cambridgeshire"
accidents$Police_Force[accidents$Police_Force ==36] <- "Norfolk"
accidents$Police_Force[accidents$Police_Force ==37] <- "Suffolk"
accidents$Police_Force[accidents$Police_Force ==40] <- "Bedfordshire"
accidents$Police_Force[accidents$Police_Force ==41] <- "Hertfordshire"
accidents$Police_Force[accidents$Police_Force ==42] <- "Essex"
accidents$Police_Force[accidents$Police_Force ==43] <- "Thames Valley"
accidents$Police_Force[accidents$Police_Force ==44] <- "Hampshire"
accidents$Police_Force[accidents$Police_Force ==45] <- "Surrey"
accidents$Police_Force[accidents$Police_Force ==46] <- "Kent"
accidents$Police_Force[accidents$Police_Force ==47] <- "Sussex"
accidents$Police_Force[accidents$Police_Force ==48] <- "City of London"
accidents$Police_Force[accidents$Police_Force ==50] <- "Devon and Cornwall"
accidents$Police_Force[accidents$Police_Force ==52] <- "Avon and Somerset"
accidents$Police_Force[accidents$Police_Force ==53] <- "Gloucestershire"
accidents$Police_Force[accidents$Police_Force ==54] <- "Wiltshire"
accidents$Police_Force[accidents$Police_Force ==55] <- "Dorset"
accidents$Police_Force[accidents$Police_Force ==60] <- "North Wales"
accidents$Police_Force[accidents$Police_Force ==61] <- "Gwent"
accidents$Police_Force[accidents$Police_Force ==62] <- "South Wales"
accidents$Police_Force[accidents$Police_Force ==63] <- "Dyfed-Powys"
accidents$Police_Force[accidents$Police_Force ==91] <- "Northern"
accidents$Police_Force[accidents$Police_Force ==92] <- "Grampian"
accidents$Police_Force[accidents$Police_Force ==93] <- "Tayside"
accidents$Police_Force[accidents$Police_Force ==94] <- "Fife"
accidents$Police_Force[accidents$Police_Force ==95] <- "Lothian and Borders"
accidents$Police_Force[accidents$Police_Force ==96] <- "Central"
accidents$Police_Force[accidents$Police_Force ==97] <- "Strathclyde"
accidents$Police_Force[accidents$Police_Force ==98] <- "Dumfries and Galloway"

accidents$Accident_Severity[accidents$Accident_Severity ==1] <- "Fatal"
accidents$Accident_Severity[accidents$Accident_Severity ==2] <- "Serious"
accidents$Accident_Severity[accidents$Accident_Severity ==3] <- "Slight"


accidents$Day_of_Week[accidents$Day_of_Week ==1] <- "Sunday"
accidents$Day_of_Week[accidents$Day_of_Week ==2] <- "Monday"
accidents$Day_of_Week[accidents$Day_of_Week ==3] <- "Tuesday"
accidents$Day_of_Week[accidents$Day_of_Week ==4] <- "Wednesday"
accidents$Day_of_Week[accidents$Day_of_Week ==5] <- "Thursday"
accidents$Day_of_Week[accidents$Day_of_Week ==6] <- "Friday"
accidents$Day_of_Week[accidents$Day_of_Week ==7] <- "Saturday"



accidents$Day_of_Week <- factor(accidents$Day_of_Week, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"), ordered=TRUE) 



```


```{r code example 3, echo = TRUE, message = FALSE}
# get counts of accidents in each police force

# get count for number of accidents in each police force and weekday
pf_accidents <- accidents %>%
  dplyr::group_by(Police_Force, Day_of_Week) %>%
  dplyr::tally() %>%
  dplyr::arrange(Police_Force, Day_of_Week) 


# ggplot - scatter for no. accidents in police force by week day

pf_scatter <- ggplot(pf_accidents, aes(Police_Force, n))+
  geom_point(aes(colour = Day_of_Week)) +
  ggtitle("Reported road accidents by Police Force and day of the week, 2017") +
  ylab("Number of Accidents in 2017") +
  scale_y_continuous(breaks = seq(0, 4400, by = 200)) +
  theme(axis.title = element_text(family = "Arial", size = 7)) +
   theme_classic() +
    theme(axis.text.x = element_text(angle = 90)) +
    scale_color_manual(values=dft_col(n=7, name="dft")) 

ggplotly(pf_scatter)

```

## create DfT colour palette


```{r}
# create dft colour palette

library(fBasics)

r <- c(0, 102, 210, 228, 201, 233, 0, 153)
g <- c(104, 164, 95, 159, 146, 211, 153, 214)
b <- c(83, 152, 21, 115, 18, 160, 169, 221)

dft_col <- function (n, name = c("dft")) 
{
  dft = rgb(r, g, b, maxColorValue = 255)
  name = match.arg(name)
  orig = eval(parse(text = name))
  rgb = t(col2rgb(orig))
  temp = matrix(NA, ncol = 3, nrow = n)
  x = seq(0, 1, , length(orig))
  xg = seq(0, 1, , n)
  for (k in 1:3) {
    hold = spline(x, rgb[, k], n = n)$y
    hold[hold < 0] = 0
    hold[hold > 255] = 255
    temp[, k] = round(hold)
  }
  palette = rgb(temp[, 1], temp[, 2], temp[, 3], maxColorValue = 255)
  palette
}

```

```{r, echo = FALSE}
# check palette colours
pal2 <- dft_col(n=8)
par(mar = rep(0, 4))
pie(rep(1, length(pal2)), col = pal2)
```

## The harder way

- Using the plot_ly() wrapper and learning a new set of arguments.

- Plotly is very flexible - double check which arguments are in which wrappers as plotly pipes arguments into each other.

- Plotly exists in R, Python, MATLAB and JaveScript. When searching for guidance on arguments, make sure it's the correct language.

- Plot_ly() over ggplotly() provides you with more options to control, including animating graphs with timelines, buttons and sliders.



```{r libraries and data import, include= FALSE}

library("pacman")

pacman::p_load("tidyverse", "dplyr", "plyr", "plotly", "odbc","DBI")

#loading in accident datset
accidents <- read.csv("acc.csv")

#rename accident index (weird hidden characters)
accidents <- dplyr::rename(accidents, Accident_Index = "ï..Accident_Index")
#accidents <- dplyr::rename(accidents, Accident_Index = "?..Accident_Index")

#make accident index factor not character
accidents$Accident_Index <- as.character(accidents$Accident_Index)

```


``` {r cleaning data for bar chart, include = FALSE, message = FALSE}
#preparing the police force variable

#corresponding values are found in the data.gov. link in the "variable lookup" under supporting documents
accidents$Police_Force[accidents$Police_Force ==1] <- "Metropolitan Police"
accidents$Police_Force[accidents$Police_Force ==3] <- "Cumbria"
accidents$Police_Force[accidents$Police_Force ==4] <- "Lancashire"
accidents$Police_Force[accidents$Police_Force ==5] <- "Merseyside"
accidents$Police_Force[accidents$Police_Force ==6] <- "Greater Manchester"
accidents$Police_Force[accidents$Police_Force ==7] <- "Cheshire"
accidents$Police_Force[accidents$Police_Force ==10] <- "Northumbria"
accidents$Police_Force[accidents$Police_Force ==11] <- "Durham"
accidents$Police_Force[accidents$Police_Force ==12] <- "North Yorkshire"
accidents$Police_Force[accidents$Police_Force ==13] <- "West Yorkshire"
accidents$Police_Force[accidents$Police_Force ==14] <- "South Yorkshire"
accidents$Police_Force[accidents$Police_Force ==16] <- "Humberside"
accidents$Police_Force[accidents$Police_Force ==17] <- "Cleveland"
accidents$Police_Force[accidents$Police_Force ==20] <- "West Midlands"
accidents$Police_Force[accidents$Police_Force ==21] <- "Staffordshire"
accidents$Police_Force[accidents$Police_Force ==22] <- "West Mercia"
accidents$Police_Force[accidents$Police_Force ==23] <- "Warwickshire"
accidents$Police_Force[accidents$Police_Force ==30] <- "Derbyshire"
accidents$Police_Force[accidents$Police_Force ==31] <- "Nottinghamshire"
accidents$Police_Force[accidents$Police_Force ==32] <- "Lincolnshire"
accidents$Police_Force[accidents$Police_Force ==33] <- "Leicestershire"
accidents$Police_Force[accidents$Police_Force ==34] <- "Northamptonshire"
accidents$Police_Force[accidents$Police_Force ==35] <- "Cambridgeshire"
accidents$Police_Force[accidents$Police_Force ==36] <- "Norfolk"
accidents$Police_Force[accidents$Police_Force ==37] <- "Suffolk"
accidents$Police_Force[accidents$Police_Force ==40] <- "Bedfordshire"
accidents$Police_Force[accidents$Police_Force ==41] <- "Hertfordshire"
accidents$Police_Force[accidents$Police_Force ==42] <- "Essex"
accidents$Police_Force[accidents$Police_Force ==43] <- "Thames Valley"
accidents$Police_Force[accidents$Police_Force ==44] <- "Hampshire"
accidents$Police_Force[accidents$Police_Force ==45] <- "Surrey"
accidents$Police_Force[accidents$Police_Force ==46] <- "Kent"
accidents$Police_Force[accidents$Police_Force ==47] <- "Sussex"
accidents$Police_Force[accidents$Police_Force ==48] <- "City of London"
accidents$Police_Force[accidents$Police_Force ==50] <- "Devon and Cornwall"
accidents$Police_Force[accidents$Police_Force ==52] <- "Avon and Somerset"
accidents$Police_Force[accidents$Police_Force ==53] <- "Gloucestershire"
accidents$Police_Force[accidents$Police_Force ==54] <- "Wiltshire"
accidents$Police_Force[accidents$Police_Force ==55] <- "Dorset"
accidents$Police_Force[accidents$Police_Force ==60] <- "North Wales"
accidents$Police_Force[accidents$Police_Force ==61] <- "Gwent"
accidents$Police_Force[accidents$Police_Force ==62] <- "South Wales"
accidents$Police_Force[accidents$Police_Force ==63] <- "Dyfed-Powys"
accidents$Police_Force[accidents$Police_Force ==91] <- "Northern"
accidents$Police_Force[accidents$Police_Force ==92] <- "Grampian"
accidents$Police_Force[accidents$Police_Force ==93] <- "Tayside"
accidents$Police_Force[accidents$Police_Force ==94] <- "Fife"
accidents$Police_Force[accidents$Police_Force ==95] <- "Lothian and Borders"
accidents$Police_Force[accidents$Police_Force ==96] <- "Central"
accidents$Police_Force[accidents$Police_Force ==97] <- "Strathclyde"
accidents$Police_Force[accidents$Police_Force ==98] <- "Dumfries and Galloway"


```

## Bar Charts in plotly 

```{r example code 4, echo = TRUE, message=FALSE}

#Grouping accidents by police force
accidentsPF <- accidents %>%          #SQL FROM clause equivalent
  dplyr::group_by(Police_Force) %>%   #SQL SELECT/GROUP BY clause equivalent
  dplyr::tally()                      #SQL count(*) equivalent



#reorder the data so that you change the police forces to graph in descending order
accidentsPF$Police_Force <- factor(accidentsPF$Police_Force, 
                                   levels = unique(accidentsPF$Police_Force)[order(accidentsPF$n, decreasing = TRUE)])


#adding more information for the bar chart
accidentsPF$percent <- (accidentsPF$n/129982)*100
accidentsPF$percent <- round(accidentsPF$percent, 1)


#creating a bar chart
accidentsPF %>%
  plot_ly(
    y = ~ Police_Force, #use a ~ to tell plot_ly that you want to reference a variable
    x = ~ n,
    type = "bar", #specifies what chart you want
    marker = list(color = 'rgb(210,95,21)'),
    hoverinfo = "text", #dialogue box that appears when you mouse over a data point
    hovertext = paste(
      "<b>Force: </b>", accidentsPF$Police_Force,
      "<i><br>Count: </i>",accidentsPF$n,
      "<br>Per cent: ", accidentsPF$percent)
    ) %>%
      layout(title = "Personal injury reported road accidents in 2017 by Police Force",
             xaxis = list(title = "Count",
                          showgrid = F,
                          zeroline = F),
             yaxis = list(title = "Police Force",
                          showgrid = F,
                          zeroline = F)
              )

```

## Scatter charts in plotly 

```{r scatter plot data import, echo = TRUE, message = FALSE}

con <- dbConnect(odbc::odbc(),
                 .connection_string = "database_server")

accidentsscatter <- dbGetQuery(con,
                              "select a.accyr as year, a1lab as policeforce,  count(distinct a.accid) as accidents, count(*) as casualties
                              from cas c left join acc a on c.accid=a.accid
                              where a.accyr between 2000 and 2017
                              group by a.accyr, a1lab
                              order by a.accyr") 

```



```{r scatter plot, echo = TRUE, MESSAGE = FALSE, warnings = FALSE}

accidentsscatter %>%
  plot_ly() %>%
  add_trace(
    x = ~ accidents, 
    y = ~ casualties, 
    color = ~ policeforce,
    frame = ~ year, 
    type = 'scatter',
    mode = 'markers',
    hoverinfo = "text",
    text = ~paste('</br><b>Force</b>: ',policeforce,'</br>Accidents: ',accidents,'</br>casualties: ',casualties)
  ) %>% animation_opts(
    1000, #animation speed in ms
    easing = "elastic", #adds "elastic animation effect 
    redraw = FALSE #TRUE will redraw plot at the completion of the transition 
  ) %>%
  layout(title = "Accidents and casualties by Police Force 2000-2018",
         xaxis = list(title = "Accidents"),
         yaxis = list(title = "Casualties")
  )


```

## Heatmap/Matrix 

``` {r heatmap data cleaning and importing, include=FALSE}

#recoding day of week variable
accidents$DayofWeek[accidents$Day_of_Week == 1] <- 'Sunday'
accidents$DayofWeek[accidents$Day_of_Week == 2] <- 'Monday'
accidents$DayofWeek[accidents$Day_of_Week == 3] <- 'Tuesday'
accidents$DayofWeek[accidents$Day_of_Week == 4] <- 'Wednesday'
accidents$DayofWeek[accidents$Day_of_Week == 5] <- 'Thursday'
accidents$DayofWeek[accidents$Day_of_Week == 6] <- 'Friday'
accidents$DayofWeek[accidents$Day_of_Week == 7] <- 'Saturday'
count(accidents$DayofWeek)

accidents$hour <- substr(accidents$Time, 1,2)
#count(accidents$hour)

accidents$hour2[accidents$hour == "00" |accidents$hour == "01" ] <- "00:00-01:59"
accidents$hour2[accidents$hour == "02" |accidents$hour == "03" ] <- "02:00-03:59"
accidents$hour2[accidents$hour == "04" |accidents$hour == "05" ] <- "04:00-05:59"
accidents$hour2[accidents$hour == "06" |accidents$hour == "07" ] <- "06:00-07:59"
accidents$hour2[accidents$hour == "08" |accidents$hour == "09" ] <- "08:00-09:59"
accidents$hour2[accidents$hour == 10 |accidents$hour == 11 ] <- "10:00-11:59"
accidents$hour2[accidents$hour == 12 |accidents$hour == 13 ] <- "12:00-13:59"
accidents$hour2[accidents$hour == 14 |accidents$hour == 15 ] <- "14:00-15:59"
accidents$hour2[accidents$hour == 16 |accidents$hour == 17 ] <- "16:00-17:59"
accidents$hour2[accidents$hour == 18 |accidents$hour == 19 ] <- "18:00-19:59"
accidents$hour2[accidents$hour == 20 |accidents$hour == 21 ] <- "20:00-21:59"
accidents$hour2[accidents$hour == 22 |accidents$hour == 23 ] <- "22:00-23:59"
count(accidents$hour2)

accidents$hour2[accidents$hour2 != "00:00-01:59"&
                accidents$hour2 != "02:00-03:59"&
                accidents$hour2 != "04:00-05:59"&
                accidents$hour2 != "06:00-07:59"&
                accidents$hour2 != "08:00-09:59"&
                accidents$hour2 != "10:00-11:59"&
                accidents$hour2 != "12:00-13:59"&
                accidents$hour2 != "14:00-15:59"&
                accidents$hour2 != "16:00-17:59"&
                accidents$hour2 != "18:00-19:59"&
                accidents$hour2 != "20:00-21:59"&
                accidents$hour2 != "22:00-23:59"
                  ] <- NA

#using lubridate to get hour out
count(accidents$Time)
#hour <- lubridate::hour(accidents$Time)    

#Grouping accidents by day of week and month
accidentsheatmap <- accidents %>%
  dplyr::group_by(DayofWeek, hour2) %>%
  dplyr::tally()



#order day of week for this chart as it is more applicable
accidentsheatmap$DayofWeek <- factor(accidentsheatmap$DayofWeek, levels=c("Sunday", "Saturday", "Friday", "Thursday", "Wednesday", "Tuesday", "Monday"), ordered=TRUE)

#ordering 

#accidentsheatmap$hour2 <- as.factor(accidentsheatmap$hour2)
```

```{r heatmap, echo = TRUE, message = FALSE}

#create dft colour scale for heatmap
palette <- colorRampPalette(c("#F4E9D0","#C99212"))


  plot_ly(
    data = accidentsheatmap,
    x = ~ hour2,
    y = ~ DayofWeek,
    z = ~ n,
    type = "heatmap",
    colors =  palette(3700)
    ) %>%
  layout(title = "Personal injury reported road accidents in 2017 by time and day of week",
         xaxis = list(title = "Time"),
         yaxis = list(title = "Day of week"),
         hoverlabel = list(font=list(size=20))
  )

```

##Any questions?