---
title: "Olympics Report 2008"
author: "Sue Wallace"
date: "09/09/2018"
output: word_document
---



```{r, include=FALSE, echo=FALSE, message=FALSE}
#load libraries----

library(readr) #reading in data and writing out data
library(dplyr) #data wrangle
library(tidyverse) #learn more about the philosophy of tidy verse here
#https://www.tidyverse.org/
library(tidyr) #cleaning a data set

#load data----

olympics <- read_csv ("Data/Summer Olympic medallists 1896 to 2008.csv") 

regions <- read_csv ("Data/noc_regions.csv")

#data wrangle----

#Let's look at the 2008 Olympics. We can join on the region data as well
#so that we know which country the athletes are from

#I'm also only really interested in a few variables

olympics %>%
  dplyr::left_join(
    x = olympics,  # to this table...
    y = regions,   # ...join this table
    by = "NOC"  # on this key
  ) %>% 
  filter(Edition == "2008") %>%
  select(Sport, Discipline, Athlete, Gender, Event,
         Medal, region) -> olympics_2008

#Which country won the most medals in 2008? Plot into a chart

olympics_2008 %>%
  group_by(region) %>% #group by the region variable
  summarise(count = n()) %>% #summarise the number in the group
  mutate(Percentage = count/sum(count)*100) %>% #create a new % variable
  top_n(10) -> medals #select the top 10 

medals %>%
  filter(region == "USA") -> usa_prop

#We can make a chart from this too

ggplot(data = medals, aes(x = reorder(region, -Percentage), y = Percentage, 
                          label = (round(Percentage,1)))) + 
  geom_col(position = "dodge", fill="lightgoldenrod1") +
  geom_text(position = position_dodge(width = .9), vjust=-0.5) +
  theme_classic() +
  ggtitle("Nation which won the most medals in the 2008 Beijing Olympics") +
  xlab("Nation") +
  ylab("Proportion") -> country


#The USA collected the highest proportion of medals. Let's do a bit more
#analysis on the states

#1. Which gender collected more medals, male or female?

olympics_2008 %>%
  filter(region == "USA") %>%
  group_by(Gender) %>%
  summarise(count=n()) -> usa_gender

usa_gender %>%
  filter(Gender == "Women") -> women

usa_gender %>%
  filter(Gender == "Men") -> men


ggplot(data = usa_gender, aes(x = reorder(Gender, -count), y = count, 
                          label = count,1))+ 
  geom_col(position = "dodge", fill = "lightgoldenrod1") +
  geom_text(position = position_dodge(width = .9), vjust=-0.5)+
  theme_classic() +
  ggtitle("Split of medals between men and women") +
  xlab("Nation") +
  ylab("Count of medals") -> gender



#2. In which event was the most medals won?

olympics_2008 %>%
  filter (region=="USA") %>%
  group_by(Discipline) %>%
  summarise(count=n()) %>%
  mutate(Percentage=count/sum(count)*100) %>%
  top_n(15) -> discipline

discipline %>%
  filter(Discipline=="Swimming") -> swimming

ggplot(data = discipline, aes(x = reorder(Discipline, -Percentage), 
                              y = Percentage, 
                          label = (round(Percentage,1)))) + 
  geom_col(position = "dodge", fill="lightgoldenrod1") +
  geom_text(position = position_dodge(width = .9), vjust=-0.5) +
  theme_classic() +
  ggtitle("Which dicipline brought home the most medals for the USA") +
  xlab("Discipline") +
  ylab("Proportion") -> discipline_chart




```

## Which nation won the most medals in the 2008 Beijing Olympics?
```{r, echo=FALSE, fig.width=6,fig.height=5}
plot(country)
```

The USA collected the highest proportion of medals at `r round(usa_prop$Percentage, 1)`%. Let's do a bit more analysis on the Unites States. 

##Which gender collected more medals, male or female?
```{r, echo=FALSE, fig.width=6,fig.height=5}
plot(gender)
```

We can see here that women collected `r women$count` medals for the US Olympics team
in the 2008 Beijing Olympics compared to `r men$count` for men.  

We've seen that more women won medals in comparison to men, but which discipline brought home the most medals for the United States?

##Which discipline brought home the most medals?
```{r, echo=FALSE, fig.width=6,fig.height=5}
plot(discipline_chart)
```
It looks like the most medals were won by the US swimming team, with `r round(swimming$Percentage, 1)`% of medals being collected within the swimming discipline.


